SeisAcoustic

### 2019-08-22 docs/latex/Acoustic_FWI/paper.tex
fix documentation

\usepackage{xy}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{url}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{color}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amssymb}      
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{booktabs}

@@

@@

With this property, the equation \ref{eq71} can be further simplified as
\frac{\partial J}{{\partial \bf m}_l} &= \sum_{it=2}^{nt} \left( \frac{{\partial \bf M}_{pz}{\bf B}_{vz}}{\partial m_l} \cdot {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it} \right)^T \tilde{{\bf p}}_z^{it+\frac{1}{2}} +  \left( \frac{{\partial \bf M}_{px}{\bf B}_{vx}}{\partial m_l} \cdot {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)^T \tilde{{\bf p}}_x^{it+\frac{1}{2}} \\
&= \sum_{it=2}^{nt} \left( \frac{{\partial \bf M}_{pz}{\bf B}_{vz}}{\partial m_l} \cdot {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it} +  \frac{{\partial \bf M}_{px}{\bf B}_{vx}}{\partial m_l} \cdot {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)^T \tilde{{\bf p}}_x^{it+\frac{1}{2}}  \\
&=\sum_{it=2}^{nt} \left( 2\rho[l]\cdot v[l] \cdot dt \cdot \left( {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it}\right)[l] +  2\rho[l]\cdot v[l] \cdot dt \cdot \left( {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)[l] \right) \cdot \tilde{{\bf p}}_x^{it+\frac{1}{2}} [l]\\
&=\sum_{it=2}^{nt} 2\rho[l]\cdot v[l] \cdot dt \cdot \left( {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it}+  {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)[l] \cdot \tilde{{\bf p}}_x^{it+\frac{1}{2}} [l] \\
\label{eq77}
The gradient of the objective function with respect to all the model parameter is given as
\frac{\partial J}{{\partial \bf m}} = \sum_{it=2}^{nt} {\bf diagm} \left( 2 {\bf \rho} \ast {\bf v} \ast dt \ast \left( {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it}+  {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)   \right) \tilde{{\bf p}}_x^{it+\frac{1}{2}} ,
\label{eq78}
where the symbol $*$ is element-wise multiplication and the operator ${\bf diagm}$ representing creating a diagonal matrix from a vector.
Summing the last two equations of equation \ref{eq46},  we can get
p[i,j,it+\frac{1}{2}] =  p[i,j,it-\frac{1}{2}] &+ k[i,j] \, dt \left(\frac{1}{dz} \sum_{n=1}^N  a_n \left( v_z[i+n-\frac{1}{2},j,it] - v_z[i-n+\frac{1}{2},j,it] \right)  \right) \\
&+ k[i,j]\, dt \left(\frac{1}{dx} \sum_{n=1}^N  a_n \left(v_x[i,j+n-\frac{1}{2},it] - v_x[i,j-n+\frac{1}{2},it] \right) \right) \\
&+ dt \cdot f[it]   
\label{eq79}
From above equations we can get the result
 \left( {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it}+  {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right) = 
 \frac{{\bf p}^{it+\frac{1}{2}} - {\bf p}^{it-\frac{1}{2}}  - dt \cdot {\bf f}^{it}}{{\bf k}\cdot dt}
 \label{eq80}
We use the expression in equation \ref{eq80} to replace the term in equation \ref{eq77}, we can get
\frac{\partial J}{{\partial \bf m}} = \sum_{it=2}^{nt} {\bf diagm} \left( \frac{2({\bf p}^{it+\frac{1}{2}} - {\bf p}^{it-\frac{1}{2}}  - dt \cdot {\bf f}^{it})}{{\bf v}} \right) \tilde{{\bf p}}_x^{it+\frac{1}{2}} 
\label{eq81}
The computations inside the ${\bf diagm}$ is element-wise operation.

@@

To simplify the derivations of gradient, we temporarily just consider one frequency slice. The synthetic data is obtained by sampling the wave field at the receiver's location, the sampling operator is indicated by operator ${\bf s}$. So the synthetic data is represented as 
\begin{equation}
{\bf d = su}
\label{eq126}
\end{equation}
In FWI, the model parameter is updated by minimizing the difference between the synthetic and observed data, here we just use the $l2$ norm as the metric to evaluate the difference, so the objective function is given as 
\begin{equation}
J = \frac{1}{2}||{\bf d} - {\bf d}_{obs}||_2^2 = \frac{1}{2}({\bf d} - {\bf d}_{obs})^T ({\bf d} - {\bf d}_{obs})^*
\label{eq127}
\end{equation}
Where ${\bf d}_{obs}$ is the observed data, the symbol $^T$ and $^*$ represent matrix transpose and complex conjugate. We assume the density is constant, so the model parameter is just the velocity model.

@@

{\bf d}^* &= {\bf s}^*{\bf u}^*

@@

J = \frac{1}{2}({\bf s}{\bf u} - {\bf d}_{obs})^T ({\bf s}{\bf u}^* - {\bf d}_{obs}^*)

@@

\frac{\partial J}{\partial {\bf u}} & = \frac{1}{2} {\bf s}^T ({\bf d}^* - {\bf d}_{obs}^*) \\
\frac{\partial J}{\partial {\bf u}^*} & = \frac{1}{2} {\bf s}^T ({\bf d} - {\bf d}_{obs}) \\

@@

\frac{\partial J}{\partial m_l} &= -\frac{1}{2} \left(    \left({\bf H}^{-1} \,\frac{\partial {\bf H}}{\partial m_l} {\bf u}\right)^T {\bf s}^T ({\bf d}^* - {\bf d}_{obs}^*)  + \left( \left({\bf H}^*\right)^{-1}  \, \frac{\partial {\bf H}^*}{\partial m_l} {\bf u}^*\right)^T  {\bf s}^T ({\bf d} - {\bf d}_{obs})   \right) \\
 &= -\frac{1}{2} \left(    \left({\bf H}^{-1} \,\frac{\partial {\bf H}}{\partial m_l} {\bf u}\right)^T {\bf s}^T ({\bf d}^* - {\bf d}_{obs}^*)  + \left( \left({\bf H}^{-1}\right)^*  \, \frac{\partial {\bf H}^*}{\partial m_l} {\bf u}^*\right)^T  {\bf s}^T ({\bf d} - {\bf d}_{obs})   \right) \\
  &= -\frac{1}{2} \left(    \left({\bf H}^{-1} \,\frac{\partial {\bf H}}{\partial m_l} {\bf u}\right)^T {\bf s}^T ({\bf d}^* - {\bf d}_{obs}^*)  + \left( \left( {\bf H}^{-1}  \, \frac{\partial {\bf H}}{\partial m_l} {\bf u} \right)^T \right)^* {\bf s}^T ({\bf d} - {\bf d}_{obs})   \right)

@@

\frac{\partial J}{\partial m_l} &= -{\mathcal R} \left\{    \left({\bf H}^{-1} \,\frac{\partial {\bf H}}{\partial m_l} {\bf u}\right)^T {\bf s}^T ({\bf d}^* - {\bf d}_{obs}^*)  \right\} \\
&= -{\mathcal R} \left\{  \left( \frac{\partial {\bf H}}{\partial m_l} {\bf u} \right)^T  ({\bf H}^T)^{-1} \,    {\bf s}^T ({\bf d}^* - {\bf d}_{obs}^*)  \right\} \\

@@

where ${\mathcal R}$ is the operator which take the real part of complex number. People usually call $\left( \frac{\partial {\bf H}}{\partial m_l} {\bf u} \right)$ as the source side wave field and $({\bf H}^T)^{-1} \,    {\bf s}^T ({\bf d}^* - {\bf d}_{obs}^*)$ as adjoint wave field, the gradient is obtained by the inner product between source-side wave field and adjoint wave field.

>>>>

\usepackage{xy}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{url}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{color}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amssymb}      
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{booktabs}
\setcounter{MaxMatrixCols}{20}

@@

The above equations are satisfied in the computation area. The adjoint pressure field at last time step  ${\tilde p}_z^{nt+\frac{1}{2}} = {\tilde p}_x^{nt+\frac{1}{2}}$ since the adjoint of sampling operator is to spray pressure recordings evenly to the two artificially splitted components. This equivalent to double the magnitude of recordings.

@@

With this property, the equation \ref{eq71} can be further simplified as
\begin{equation}
\begin{split}
\frac{\partial J}{{\partial m}_l} &= \sum_{it=2}^{nt} \left( \frac{{\partial \bf M}_{pz}{\bf B}_{vz}}{\partial m_l} \cdot {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it} \right)^T \tilde{{\bf p}}_z^{it+\frac{1}{2}} +  \left( \frac{{\partial \bf M}_{px}{\bf B}_{vx}}{\partial m_l} \cdot {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)^T \tilde{{\bf p}}_x^{it+\frac{1}{2}} \\
&= \sum_{it=2}^{nt} \left( \frac{{\partial \bf M}_{pz}{\bf B}_{vz}}{\partial m_l} \cdot {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it} +  \frac{{\partial \bf M}_{px}{\bf B}_{vx}}{\partial m_l} \cdot {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)^T \tilde{{\bf p}}_x^{it+\frac{1}{2}}  \\
&=\sum_{it=2}^{nt} \left( 2\rho[l]\cdot v[l] \cdot dt \cdot \left( {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it}\right)[l] +  2\rho[l]\cdot v[l] \cdot dt \cdot \left( {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)[l] \right) \cdot \tilde{{\bf p}}_x^{it+\frac{1}{2}} [l]\\
&=\sum_{it=2}^{nt} 2\rho[l]\cdot v[l] \cdot dt \cdot \left( {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it}+  {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)[l] \cdot \tilde{{\bf p}}_x^{it+\frac{1}{2}} [l] \\
\end{split}
\label{eq77}
\end{equation}
The gradient of the objective function with respect to all the model parameter is given as
\begin{equation}
\frac{\partial J}{{\partial \bf m}} = \sum_{it=2}^{nt} {\bf diagm} \left( 2 {\bf \rho} \ast {\bf v} \ast dt \ast \left( {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it}+  {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right)   \right) \tilde{{\bf p}}_x^{it+\frac{1}{2}} ,
\label{eq78}
\end{equation}
where the symbol $*$ is element-wise multiplication and the operator ${\bf diagm}$ representing creating a diagonal matrix from a vector.
Summing the last two equations of equation \ref{eq46},  we can get
\begin{equation}
\begin{split}
p[i,j,it+\frac{1}{2}] =  p[i,j,it-\frac{1}{2}] &+ k[i,j] \, dt \left(\frac{1}{dz} \sum_{n=1}^N  a_n \left( v_z[i+n-\frac{1}{2},j,it] - v_z[i-n+\frac{1}{2},j,it] \right)  \right) \\
&+ k[i,j]\, dt \left(\frac{1}{dx} \sum_{n=1}^N  a_n \left(v_x[i,j+n-\frac{1}{2},it] - v_x[i,j-n+\frac{1}{2},it] \right) \right) \\
&+ dt \cdot f[it]   
\end{split}
\label{eq79}
\end{equation}
From above equations we can get the result
\begin{equation}
 \left( {\bf d}_v{\bf d}_z \cdot {\bf v}_z^{it}+  {\bf d}_v{\bf d}_x \cdot {\bf v}_x^{it} \right) = 
 \frac{{\bf p}^{it+\frac{1}{2}} - {\bf p}^{it-\frac{1}{2}}  - dt \cdot {\bf f}^{it}}{{\bf k}\cdot dt}
 \label{eq80}
\end{equation}
We use the expression in equation \ref{eq80} to replace the term in equation \ref{eq77}, we can get
\begin{equation}
\frac{\partial J}{{\partial \bf m}} = \sum_{it=2}^{nt} {\bf diagm} \left( \frac{2({\bf p}^{it+\frac{1}{2}} - {\bf p}^{it-\frac{1}{2}}  - dt \cdot {\bf f}^{it})}{{\bf v}} \right) \tilde{{\bf p}}_x^{it+\frac{1}{2}} 
\label{eq81}
\end{equation}
The computations inside the ${\bf diagm}$ is element-wise operation.

@@

To simplify the derivations of gradient, we temporarily just consider one frequency slice. The synthetic data is obtained by sampling the wave field at the receiver's location, the sampling operator is indicated by operator ${\bf S}$. So the synthetic data is represented as 
\begin{equation}
{\bf d = Su}
\label{eq126}
\end{equation}
In FWI, the model parameter is updated by minimizing the difference between the synthetic and observed data, here we just use the $l2$ norm as the metric to evaluate the difference, so the objective function is given as 
\begin{equation}
J = \frac{1}{2}||{\bf d} - {\bf d}_{obs}||_2^2 = \frac{1}{2}({\bf d} - {\bf d}_{obs})^T ({\bf d} - {\bf d}_{obs})^*
\label{eq127}
\end{equation}
Where ${\bf d}_{obs}$ is the observed data, the symbol $^T$ and $^*$ represent matrix transpose and complex conjugate. We assume the density is constant, so the model parameter is just the velocity model.

@@

{\bf d}^* &= {\bf S}^*{\bf u}^*

@@

J = \frac{1}{2}({\bf S}{\bf u} - {\bf d}_{obs})^T ({\bf S}{\bf u}^* - {\bf d}_{obs}^*)

@@

\frac{\partial J}{\partial {\bf u}} & = \frac{1}{2} {\bf S}^T ({\bf d}^* - {\bf d}_{obs}^*) \\
\frac{\partial J}{\partial {\bf u}^*} & = \frac{1}{2} {\bf S}^T ({\bf d} - {\bf d}_{obs}) \\

@@

\frac{\partial J}{\partial m_l} &= -\frac{1}{2} \left(    \left({\bf H}^{-1} \,\frac{\partial {\bf H}}{\partial m_l} {\bf u}\right)^T {\bf S}^T ({\bf d}^* - {\bf d}_{obs}^*)  + \left( \left({\bf H}^*\right)^{-1}  \, \frac{\partial {\bf H}^*}{\partial m_l} {\bf u}^*\right)^T  {\bf S}^T ({\bf d} - {\bf d}_{obs})   \right) \\
 &= -\frac{1}{2} \left(    \left({\bf H}^{-1} \,\frac{\partial {\bf H}}{\partial m_l} {\bf u}\right)^T {\bf S}^T ({\bf d}^* - {\bf d}_{obs}^*)  + \left( \left({\bf H}^{-1}\right)^*  \, \frac{\partial {\bf H}^*}{\partial m_l} {\bf u}^*\right)^T  {\bf S}^T ({\bf d} - {\bf d}_{obs})   \right) \\
  &= -\frac{1}{2} \left(    \left({\bf H}^{-1} \,\frac{\partial {\bf H}}{\partial m_l} {\bf u}\right)^T {\bf S}^T ({\bf d}^* - {\bf d}_{obs}^*)  + \left( \left( {\bf H}^{-1}  \, \frac{\partial {\bf H}}{\partial m_l} {\bf u} \right)^T \right)^* {\bf S}^T ({\bf d} - {\bf d}_{obs})   \right)

@@

\frac{\partial J}{\partial m_l} &= -{\mathcal R} \left\{    \left({\bf H}^{-1} \,\frac{\partial {\bf H}}{\partial m_l} {\bf u}\right)^T {\bf S}^T ({\bf d}^* - {\bf d}_{obs}^*)  \right\} \\
&= -{\mathcal R} \left\{  \left( \frac{\partial {\bf H}}{\partial m_l} {\bf u} \right)^T  ({\bf H}^T)^{-1} \,    {\bf S}^T ({\bf d}^* - {\bf d}_{obs}^*)  \right\} \\

@@

where ${\mathcal R}$ is the operator which take the real part of complex number. People usually call $\left( \frac{\partial {\bf H}}{\partial m_l} {\bf u} \right)$ as the source side wave field and $({\bf H}^T)^{-1} \,    {\bf S}^T ({\bf d}^* - {\bf d}_{obs}^*)$ as adjoint wave field, the gradient is obtained by the inner product between source-side wave field and adjoint wave field.

We considering the details of $\frac{\partial {\bf H}}{\partial m_l}$, we temporarily only consider the velocity as the model parameter and consider density as constant (not updated by inversion). To simplify the derivation, we assume the model size (include PML) is $3 \times 4$.
The mapping between sub-index and linear index are expressed as follow
\begin{equation}
\begin{pmatrix}
[1,1] \rightarrow 1 & [1,2] \rightarrow 4 & [1,3]\rightarrow 7 & [1,4]\rightarrow 10 \\
[2,1]\rightarrow 2 & [2,2]\rightarrow 5   & [2,3]\rightarrow 8 & [2,4]\rightarrow 11 \\
[3,1]\rightarrow 3 & [3,2] \rightarrow 6  & [3,3]\rightarrow 9 & [3,4]\rightarrow 12 \\
\end{pmatrix}
\end{equation}
We only consider the velocity part and define $c_1 = -c\, \omega^2$, $c_2 = - d\,\omega^2$, $c_3 = - \frac{1-c-4d}{4}\,\omega^2$
 So the part of the Hessian matrix only related to velocity model can be expressed as
\begin{equation}
\begin{pmatrix}
\frac{c_1}{k_1} & \frac{c_2}{k_2} & \quad & \frac{c_2}{k_4} & \frac{c_3}{k_5} & \quad & \quad & \quad & \quad & \quad & \quad & \quad \\

\frac{c_2}{k_1} & \frac{c_1}{k_2} & \frac{c_2}{k_3} & \frac{c_3}{k_4} & \frac{c_2}{k_5} & \frac{c_3}{k_6}  & \quad & \quad & \quad & \quad & \quad & \quad \\

\quad & \frac{c_2}{k_2} & \frac{c_1}{k_3} & \quad & \frac{c_3}{k_5} & \frac{c_2}{k_6}  & \quad & \quad & \quad & \quad & \quad & \quad \\\\

\frac{c_2}{k_1} & \frac{c_3}{k_2}  & \quad & \frac{c_1}{k_4}  & \frac{c_2}{k_5} & \quad   & \frac{c_2}{k_7} & \frac{c_3}{k_8} & \quad & \quad & \quad & \quad \\

\frac{c_3}{k_1} & \frac{c_2}{k_2} & \frac{c_3}{k_3} & \frac{c_2}{k_4} & \frac{c_1}{k_5} & \frac{c_2}{k_6} & \frac{c_3}{k_7} & \frac{c_2}{k_8} & \frac{c_3}{k_9} & \quad & \quad & \quad \\

\quad & \frac{c_3}{k_2} & \frac{c_2}{k_3} &  \quad & \frac{c_2}{k_5} & \frac{c_1}{k_6} &  \quad & \frac{c_3}{k_8} & \frac{c_2}{k_9} &  \quad &  \quad &  \quad \\\\

\quad & \quad &\quad & \frac{c_2}{k_4]} & \frac{c_3}{k_5} & \quad & \frac{c_1}{k_7} & \frac{c_2}{k_8} & \quad & \frac{c_2}{k_{10}} & \frac{c_3}{k_{11}} & \quad \\

\quad  & \quad  & \quad & \frac{c_3}{k_4} & \frac{c_2}{k_5} & \frac{c_3}{k_6} & \frac{c_2}{k_7}  & \frac{c_1}{k_8} & \frac{c_2}{k_9}  & \frac{c_3}{k_{10}} & \frac{c_2}{k_{11}} & \frac{c_3}{k_{12}} \\

\quad &  \quad &  \quad &  \quad & \frac{c_3}{k_5} & \frac{c_2}{k_6} &  \quad & \frac{c_2}{k_8} & \frac{c_1}{k_9} &  \quad & \frac{c_3}{k_{11}} & \frac{c_2}{k_{12}} \\\\

\quad &  \quad &  \quad & \quad & \quad & \quad & \frac{c_2}{k_7} & \frac{c_3}{k_8} & \quad & \frac{c_1}{k_{10}} & \frac{c_2}{k_{11}} & \quad \\

\quad &  \quad &  \quad & \quad & \quad & \quad & \frac{c_3}{k_7} & \frac{c_2}{k_8} & \frac{c3}{k_9} & \frac{c_2}{k_{10}} & \frac{c_1}{k_{11}} & \frac{c_2}{k_{12}} \\

\quad &  \quad &  \quad & \quad & \quad & \quad & \quad & \frac{c_3}{k[_8} & \frac{c_2}{k_9} & \quad & \frac{c_2}{k_{11}} & \frac{c_1}{k_{12}} 
\end{pmatrix}
\end{equation}
We know $k = \rho \times v^2$, so $\frac{\partial \frac{c_i}{k}}{\partial\, v} = -\frac{2\, c_i \rho v}{k^2} =  -\frac{2\, c_i}{\rho v^3}$
Suppose we want to compute $\frac{\partial {\bf H}}{\partial v_5}$ is given as
\begin{equation}
\begin{pmatrix}
0 & 0 & 0 & 0 & -\frac{2\, c_3}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_2}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_3}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_2}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_1}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_2}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_3}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_2}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_3}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
\end{pmatrix}
\end{equation}
So the result of $\left( \frac{\partial {\bf H}}{\partial v_5} {\bf u} \right)$ is given as
\begin{equation}
\begin{pmatrix}
0 & 0 & 0 & 0 & -\frac{2\, c_3}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_2}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_3}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_2}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_1}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_2}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_3}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_2}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & -\frac{2\, c_3}{\rho_5 v_5^3} & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
\end{pmatrix}
\begin{bmatrix}
u_1\\
u_2\\
u_3\\
u_4\\
u_5\\
u_6\\
u_7\\
u_8\\
u_9\\
u_{10}\\
u_{11}\\
u_{12}\\
\end{bmatrix}
=
\begin{bmatrix}
-\frac{2\, c_3\, u_5}{\rho_5 v_5^3} \\
-\frac{2\, c_2\, u_5}{\rho_5 v_5^3} \\
-\frac{2\, c_3\, u_5}{\rho_5 v_5^3} \\
-\frac{2\, c_2\, u_5}{\rho_5 v_5^3} \\
-\frac{2\, c_1\, u_5}{\rho_5 v_5^3} \\
-\frac{2\, c_2\, u_5}{\rho_5 v_5^3}  \\
-\frac{2\, c_3\, u_5}{\rho_5 v_5^3} \\
-\frac{2\, c_2\, u_5}{\rho_5 v_5^3}  \\
-\frac{2\, c_3\, u_5}{\rho_5 v_5^3} \\
0 \\
0 \\
0 \\
\end{bmatrix}
=
-\frac{2\, u_5}{\rho_5 \,v_5^3}
\begin{bmatrix}
c_3 \\
c_2 \\
c_3 \\
c_2 \\
c_1 \\
c_2 \\
c_3 \\
c_2 \\
c_3 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{equation}
We also name the adjoint wave field is given as
\begin{equation}
{\bf u}_{adj} = ({\bf H}^T)^{-1} \,    {\bf S}^T ({\bf d}^* - {\bf d}_{obs}^*)
\end{equation}
The final formulation of the gradient of $v_5$ is given as
\begin{equation}
\begin{split}
g_5 &= \left( \frac{\partial {\bf H}}{\partial v_5} {\bf u} \right)^T \tilde{\bf u} \\
       &= -\frac{2\, u_5}{\rho_5 \,v_5^3} \begin{bmatrix}
       c_3 &
c_2 &
c_3 &
c_2 &
c_1 &
c_2 &
c_3 &
c_2 &
c_3 &
0 &
0 &
0 
\end{bmatrix}
\begin{bmatrix}
\tilde{u}_1 \\
\tilde{u}_2 \\
\tilde{u}_3 \\
\tilde{u}_4 \\
\tilde{u}_5 \\
\tilde{u}_6 \\
\tilde{u}_7 \\
\tilde{u}_8 \\
\tilde{u}_9 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{split}
\end{equation}
